\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2014-01-28 delays.tex

The aforementioned technique maintains the essence of synchronous and
GALS programs -- the logical tick, but in doing so is unable to provide
the exact real-time wait of $M$ time units in case of
\texttt{wait\_exact(M)} statement, thereby violating its
semantics. The second option is to execute the CD periodically at the
WCRT. WCRT determines the fastest execution speed of any CD with the
guarantee that no input events from the environment will be missed
(cf. Section~\ref{sec:mapping-logical-time}). Thus, if we simply execute
the CD at periodically at the WCRT, we can choose the value of
\textit{d} from set $S1$. This technique allows meeting the synchronous
hypothesis, while at the same time guaranteeing the semantics of the
\texttt{wait\_exact(M)} statement. But, compared to the previous
option, this technique is very inflexible and results in reduced
response time and increased power consumption.

@@

The interaction of the low-level execution platform and the described
high-level technique of introducing real-time waits in the SystemJ
language is shown in Figure~\ref{fig:4}. We use static-analysis
techniques to find out the WCRT/BCRT of the programs automatically after
rewrites and standard compiler optimizations. We require timing
analyzable platforms, in our case \textit{Java Optimized Processor}
(JOP)~\cite{jop:jnl:jsa2007}. Moreover, since algorithms~\ref{alg:1}
and~\ref{alg:2} are performed after scheduling for different constraints
such as faster execution time, or reduced power consumption, we,
inherently satisfy the criteria of exploiting non-maximal
parallelism. The low-level analysis guarantees a precise WCRT/BCRT for
the logical tick, which is then used for calculating $d$ in
Figure~\ref{fig:3}.

>>>>

The aforementioned technique maintains the essence of synchronous and
GALS programs -- the logical tick, but in doing so is unable to provide
the exact real-time wait of $M$ time units in case of
\texttt{wait\_exact(M)} statement, thereby violating its
semantics. The second option is to execute the CD periodically at the
WCRT. WCRT determines the fastest execution speed of any CD with the
guarantee that no input events from the environment will be missed
(cf. Section~\ref{sec:mapping-logical-time}). Thus, if we simply execute
the CD periodically at the WCRT, we can choose the value of
\textit{d} from set $S1$. This technique allows meeting the synchronous
hypothesis, while at the same time guaranteeing the semantics of the
\texttt{wait\_exact(M)} statement. But, compared to the previous
option, this technique is very inflexible and results in reduced
response time and increased power consumption.

@@

The interaction of the low-level execution platform and the described
high-level technique of introducing real-time waits in the SystemJ
language is shown in Figure~\ref{fig:4}. We use static analysis
techniques to find out the WCRT/BCRT of the programs automatically after
rewrites and standard compiler optimizations. We require timing
analyzable platforms, in our case \textit{Java Optimized Processor}
(JOP)~\cite{jop:jnl:jsa2007}. Moreover, since algorithms~\ref{alg:1}
and~\ref{alg:2} are performed after scheduling for different constraints
such as faster execution time, or reduced power consumption, we
inherently satisfy the criteria of exploiting non-maximal
parallelism. The low-level analysis guarantees a precise WCRT/BCRT for
the logical tick, which is then used for calculating $d$ in
Figure~\ref{fig:3}.

### 2014-01-28 discussion.tex

\subsection{Programming using the wait\_exactly construct -- the
  periodic reactions}
\label{sec:progr-using-delay}

@@

\begin{figure}[h!]
  \centering
	\vspace{-10pt}
        \begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exactly}]
          while(true){wait_exactly (1 ms); emit S; //do something
          }
 \end{lstlisting}
  \caption{Programming periodic tasks with wait\_exactly statement}
  \label{fig:periodic}
	\vspace{-10pt}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \begin{SubFloat}{\label{pp:a}Original program}
    \centering
			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exactly}]
while(true) {
 wait_exactly(1 ms);
 //extra pause
 pause;
 emit S;
}
\end{lstlisting}
\end{SubFloat}
\hspace{9pt}
  \begin{SubFloat}{\label{pp:b}Rewritten program}
			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay}]
while(true) {
 while(true) { 
  trap(T) {
   int x = 0;
   while(true) {
    x = x + 1;
    pause;
    if (x == d)
     exit(T);
   }
  }
  //extra pause
  pause;
  emit S;
 }
}
\end{lstlisting}
  \end{SubFloat}
\caption{Extra pause statements in periodic tasks}
  \label{fig:periodic2}
\end{figure}

>>>>

\subsection{Programming using the wait\_exact construct -- the
  periodic reactions}
\label{sec:progr-using-delay}

@@

\begin{figure}[h!]
  \centering
	\vspace{-10pt}
        \begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exact}]
          while(true){wait_exact (1 ms); emit S; //do something
          }
 \end{lstlisting}
  \caption{Programming periodic tasks with wait\_exact statement}
  \label{fig:periodic}
	\vspace{-10pt}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \begin{SubFloat}{\label{pp:a}Original program}
    \centering
			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,wait_exact}]
while(true) {
 wait_exact(1 ms);
 //extra pause
 pause;
 emit S;
}
\end{lstlisting}
\end{SubFloat}
\hspace{9pt}
  \begin{SubFloat}{\label{pp:b}Rewritten program}
			\begin{lstlisting}[style=sysj,basicstyle=\normalsize\ttfamily,morekeywords={emit,trap,pause,exit,delay}]
while(true) {
 while(true) { 
  trap(T) {
   int x = 0;
   while(true) {
    x = x + 1;
    pause;
    if (x == d)
     exit(T);
   }
  }
  //extra pause
  pause;
  emit S;
 }
}
\end{lstlisting}
  \end{SubFloat}
\caption{Extra pause statements in periodic tasks}
  \label{fig:periodic2}
\end{figure}

### 2014-01-28 introduction.tex

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc. Usually these systems also need to meet
real-time constraints imposed by the environment. Yet, these languages
do not support describing real-time statements as first class language
constructs.  For example, one cannot describe a simple real-time delay
(postponement) of an operation for 0.2 ms. These languages are based on
formal semantics, essential for the formal reasoning about and
verification of correctness of functional properties of the developed
programs, but leave nonfunctional properties such as timing behavior as
an implementation detail~\cite{boldt07}. Arguably, rightly so, because
physical time cannot be incorporated without information about the
underlying execution platform.  But, one can still reason in terms of
logical time. These languages support discrete logical clock rather than
a discrete physical (real-time) clock. The real-time period of the
discrete logical clock (also referred to as a logical tick) is not fixed
and it is determined by the responsiveness of the program to the input
signals. Unlike a discrete physical clock, which has fixed real-time
period, the period of the logical clock is elastic. Although the timing
model with logical clock has worked well for reactive synchronous/GALS
languages in designing discrete systems, there is a need to introduce
real-time, since often, implementation models are in the continuous or
discrete real-time domain~\cite{DBLP:journals/pieee/SifakisTY03}.
Moreover, adding timing capabilities as first class language constructs
create additional opportunity for formally verifying functional and
non-functional requirements before system deployment.

@@

The major motivations for introducing these real-time \textit{wait}
mechanisms and at the same time \textbf{contributions} of the paper are:
(1) these waiting mechanisms allow real-time based synchronization
between concurrent behaviors, (2) the waiting model relies on the use of
relative instead of absolute real-time, i.e., a wait in selected time
units is counted from the currently executing SystemJ statement, (3)
these waiting mechanisms allow mixing behaviors with real-time features
and others with only logical time and (4) real-time wait does not affect
the model of logical time in SystemJ program as long as the amount of
delay is within boundaries that can be determined statically by program
analysis. All these as well as semantics of the wait constructs are
discussed and illustrated in the rest of the paper. Our contributions
can be further refined as follows:
\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
 i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reducing power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time waits.
\item We are the first, to our knowledge, to allow specification of
  both; exact and non-exact real-time in GALS and synchronous programs.
\item We elucidate upon a key insight relating real-time and
  logical-time -- \textit{the resolution of any real-time statement is
    dependent on the worst case logical tick time} -- an aspect that
  seems to have escaped in the current literature on incorporation of
  real-time in synchronous/GALS languages.
\end{enumerate*}

>>>>

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc. Usually these systems also need to meet
real-time constraints imposed by the environment. Yet, these languages
do not support describing real-time statements as first class language
constructs.  For example, one cannot describe a simple real-time delay
(postponement) of an operation for 0.2 ms. These languages are based on
formal semantics, essential for the formal reasoning about \hj{code
generation} and verification of correctness of functional properties of
the developed programs, but leave nonfunctional properties such as
timing behavior as an implementation detail~\cite{boldt07}. Arguably,
rightly so, because physical time cannot be incorporated without
information about the underlying execution platform.  But, one can still
reason in terms of logical time. These languages support discrete
logical clock rather than a discrete physical (real-time) clock. The
real-time period of the discrete logical clock (also referred to as a
logical tick) is not fixed and it is determined by the responsiveness of
the program to the input signals. Unlike a discrete physical clock,
which has fixed real-time period, the period of the logical clock is
elastic. Although the timing model with logical clock has worked well
for reactive synchronous/GALS languages in designing discrete systems,
there is a need to introduce real-time since often, implementation
models are in the continuous or discrete real-time
domain~\cite{DBLP:journals/pieee/SifakisTY03}.  Moreover, adding timing
capabilities as first class language constructs create additional
opportunity for formally verifying functional and non-functional
requirements before system deployment.

@@

The major motivations for introducing these real-time \textit{wait}
mechanisms and at the same time \textbf{contributions} of the paper are:
\hj{(1) real-time wait does not affect the model of logical time in
	SystemJ program as long as the amount of delay is within boundaries
	that can be determined statically by program analysis,} (2) the
	waiting model relies on the use of relative instead of absolute
	real-time, i.e., a wait in selected time units is counted from the
	currently executing SystemJ statement and (3) these waiting mechanisms
	allow mixing behaviors with real-time features and others with only
	logical time. All these as well as semantics of the wait constructs
	are discussed and illustrated in the rest of the paper. Our
	contributions can be further refined as follows:

\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
 i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reducing power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time waits.
\item We are the first, to our knowledge, to allow specification of
  both; exact and non-exact real-time in GALS and synchronous programs.
\item We \hj{clarify} upon a key insight relating real-time and
  logical-time -- \textit{the resolution of any real-time statement is
    dependent on the worst case logical tick time} -- an aspect that
  seems to have escaped in the current literature on incorporation of
  real-time in synchronous/GALS languages.
\end{enumerate*}
